### 2016-08-25 notions.tex
updated privacy defs

\subsection{Notions of Privacy}
One can also consider the orthogonal notion of privacy for a
set-representation.  We formalize two types of privacy notions. One
type captures a kind of one-wayness of the representation~$M$,
relative to the allowed queries.  Informally, that given the
representation and the public information, it is hard to guess any
value in the represented set~$S$.  The other type captures a kind of
semantic security, essentially requiring that $(M,\pub)$ and th
allowed queries do not leak any (non-trivial) information about the set~$S$.
%\jnote{Actually, I am now unsure what the right way to define it is. It seems natural
%to give the attacker $\aux$. But if we give them $\aux$ and $M$ then they have everything.
%If $S$ has high min-entropy (element-wise), then it might still be
%possible to have security here.}\tsnote{I think we will always bump up
%against the (element-wise) min-entropy of the distribution over~$S$,
%at least in the ``one-wayness'' style definition that I had been imagining. }

In Figure~\ref{fig:privacy-ow} we give two ``one-wayness''
definitions, in which the adversary must output an element of the
set~$S$ represented by~$M$.  The first version (priv) gives the
adversary~$(M,\pubaux)$ as input; the second (wpriv) gives
only~$\pubaux$.  The wpriv-notion captures leakage about the elements
of~$S$ that results from observing the results of allowed queries.
The stronger priv-notion captures what is jointly leaked, by the
representation~$M$ and the allowed queries, about set elements.  We
define advantage measures $\AdvPriv{\Pi,\distr{}{}}{A} =
\Prob{\ExpPriv{\Pi,\distr{}{}}{A}=1}$ and $\AdvWPriv{\Pi,\distr{}{}}{A} =
\Prob{\ExpWPriv{\Pi,\distr{}{}}{A}=1}$; in both we track the
time-complexity~$t$ and number of oracle queries~$q$ made by the adversary.

\tsnote{Wet paint...}
In Figure~\ref{fig:privacy-ss} we give a definition capturing semantic-security (of sets)
with respect to the full representation (ss-rep).  In the ``real''
setting (ss-rep-1), a set~$S$ is sampled according to distribution~$\distr{}{}$,
and the adversary is given the~$M,\pubaux$ that result from running
$\Rep(S)$.  It is also provided oracle access to $\TestOracle$.
The adversary must output a function~$f\colon\calS \to \bits^*$ and a
string~$v$ such that $f(S)=v$.  In the ``fake'' setting (ss-rep-0) a
predictor~$P$ is given $\pubaux$ only, and no $\TestOracle$ access.  It
also outputs an $f,v$ pair such that $f(S)=v$.  The advantage of~$A$ (over~$P$)
is $\AdvPrivSS{\Pi,\distr{}{}}{A,P} =
\Prob{\ExpPrivSSreal{\Pi,\distr{}{}}{A}=1} -
\Prob{\ExpPrivSSsim{\Pi,\distr{}{}}{P}=1}$.  To prevent trivial wins
for~$A$, e.g. $f(S)=M$, we forbid~$A$ from outputting~$f$
such that $f(S)=f(S')$ for all $S,S'$ in the support of the
distribution $\distr{}{}$.

\begin{figure}[htbp]
\centering
\fpage{.495}{
\hpagess{.5}{.45}
{
\experimentv{$\ExpPriv{\setprim,\distr{}{}}{A}$}\\
$S \getsr \distr{}{}$\\
$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
$z \getsr A^{\TestOracle}(M,\pubaux)$\\
if $z \in S$ then Return 1\\
Return 0\\

\medskip
\experimentv{$\ExpWPriv{\setprim,\distr{}{}}{A}$}\\
$S \getsr \distr{}{}$\\
$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
$z \getsr A^{\TestOracle}(\pubaux)$\\
if $z \in S$ then Return 1\\
Return 0\\
}
%
{
\oracle{$\TestOracle(q)$}\\
$a \gets \Qry(M,\pubaux,\privaux,q)$\\
Return~$a$
}
}
\caption{``One-wayness'' style privacy
  definitions. The bottom is a weaker notion, because the representation
  is hidden from the adversary.} 
\label{fig:privacy-ow}
\end{figure}

%
\begin{figure}[htbp]
\centering
\hfpages{.495}
{
\hpagess{.49}{.49}
{

\experimentv{$\ExpPrivSSreal{\setprim,\distr{}{}}{A}$}\\
$S\getsr \distr{}{}$\\
$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
$(f,v) \getsr A^{\TestOracle}(M,\pubaux)$\\
if $f(S)=v$ then\\
\nudge Return 1\\
Return 0\\

\medskip
\experimentv{$\ExpPrivSSsim{\setprim,\distr{}{}}{P}$}\\
$S\getsr \distr{}{}$\\
$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
$(f,v) \getsr P(\pubaux)$\\
if $f(S)=v$ then\\
\nudge Return 1\\
Return 0\\
}
%
{
\oracle{$\TestOracle(q)$}\\
$a \gets \Qry(M,\pubaux,\privaux,q)$\\
Return~$a$\\
}
}
{
\hpagess{.49}{.49}
{
\experimentv{$\ExpPrivSS{\setprim,\distr{}{},P}{A}$}\\
$S\getsr \distr{}{}$\\
$b \getsr \bits$\\
%$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
$(M_1,\pubaux,\privaux) \getsr \Rep(S)$\\
$M_0 \getsr P(\pubaux)$\\
%$(f,v) \getsr \getsr A^{\TestOracle}(M,\pubaux)$\\
$b' \getsr A^{\TestOracle}(M_b,\pubaux)$\\
%if $f(S)=v$ then\\
if $b'=b$ then\\
\nudge Return 1\\
Return 0\\
%
%\medskip
%\experimentv{$\ExpPrivSSsim{\setprim,\distr{}{}}{P}$}\\
%$S\getsr \distr{}{}$\\
%$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
%$(f,v) \getsr P(\pubaux)$\\
%if $f(S)=v$ then\\
%\nudge Return 1\\
%Return 0\\
}
%
{
\oracle{$\TestOracle(q)$}\\
$a \gets \Qry(M_1,\pubaux,\privaux,q)$\\
Return~$a$\\
}
}
\caption{Semantic-security style privacy
  definitions. \textcolor{red}{On the left, need to protect against
    adversary outputting
    things like $f(S)=M$.  One idea is to forbid~$f$ such that
    $f(S)=f(S')$ for all $S,S'$ in the support of $\distr{}{}$.}
    } 
\label{fig:privacy-ss}
\end{figure}

>>>>

\subsection{Notions of Privacy}
One can also consider the orthogonal notion of privacy for a
data structure. \jnote{I am still hoping that we can treat data structures
at an even more general level than set-representation data structures.}
We formalize two types of privacy notions. The first captures
a ``one-wayness'' property, which (informally) requires that an attacker
cannot learn any value in a set $S$ given its
representation $M$ along with the public information~$\pub$.  The second definition
captures a flavor of semantic security, essentially requiring that $M$ and $\pub$ do
not leak any (non-trivial) information about the set~$S$ they represent.
%\jnote{Actually, I am now unsure what the right way to define it is. It seems natural
%to give the attacker $\aux$. But if we give them $\aux$ and $M$ then they have everything.
%If $S$ has high min-entropy (element-wise), then it might still be
%possible to have security here.}\tsnote{I think we will always bump up
%against the (element-wise) min-entropy of the distribution over~$S$,
%at least in the ``one-wayness'' style definition that I had been imagining. }

\jnote{OK, from here on I specialize to set-representation data structures\ldots}
The one-wayness definition (see
Figure~\ref{fig:privacy-ow}) is parameterized by a distribution $\distr{}{}$ over
sets. \jnote{I got rid of the weaker definition, since it seems unnatural to me.}
A set $S$ is chosen from this distribution, and a representation
$(M,\pubaux,\privaux)$ of that set is computed.
In one experiment ($\ExpPriv{\Pi,\distr{}{}}$), the attacker is given $M, \pubaux$;
in a second experiment (${\bf Exp}^{\rm null}_{\Pi,\distr{}{}}$), it is not.
In both cases, the attacker is given
access to an oracle via which it can
issue queries and obtain the corresponding responses.
For a given $\distr{}{}$ and $A$, we define the advantage measure
$\AdvPriv{\Pi,\distr{}{}}{A} \bydef
\left| \Prob{\ExpPriv{\Pi,\distr{}{}}{A}=1} - \Prob{{\bf Exp}^{\rm null}_{\Pi,\distr{}{}}(A)=1} \right|$.
We track the
time-complexity~$t$ and number of oracle queries~$q$ made by the adversary.

\begin{figure}[hbtp]
\centering
\fpage{.6}{
\hpagess{.5}{.45}
{
\experimentv{$\ExpPriv{\setprim,\distr{}{}}{A}$}\\
$S \getsr \distr{}{}$\\
$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
$z \getsr A^{\TestOracle}(M,\pubaux)$\\
if $z \in S$ then return 1\\
return 0\\

\medskip
\experimentv{${\bf Exp}^{\rm null}_{\Pi, \distr{}{}}(A)$} \\
$S \getsr \distr{}{}$\\
$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
$z \getsr A^{\TestOracle}()$\\
if $z \in S$ then return 1\\
return 0\\
%\medskip
%\experimentv{$\ExpWPriv{\setprim,\distr{}{}}{A}$}\\
%$S \getsr \distr{}{}$\\
%$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
%$z \getsr A^{\TestOracle}(\pubaux)$\\
%if $z \in S$ then Return 1\\
%Return 0\\
}
%
{
\oracle{$\TestOracle(q)$}\\
$a \gets \Qry(M,\pubaux,\privaux,q)$\\
return~$a$
}
}
\caption{Experiments for a ``one-wayness'' style privacy
  definition.}
\label{fig:privacy-ow}
\end{figure}

\jnote{I'm not sure whether we want to limit consideration to ``high-entropy''
distributions
and require the probability of finding an element to be low,
or whether we want to allow arbitrary distributions but compare the attacker's
success probability when given $(M, \pub)$ to its success probability when given nothing.
The latter seems cleaner, so I took that approach here.}

\jnote{Actually, I am unsure whether a one-wayness definition is independently
interesting if we have (and can achieve) a semantic-security definition.}

Our semantic-secure definition is simulation-based (see
Figure~\ref{fig:privacy-ss}).
In this definition we compare the probabilities with which an attacker $A$ outputs~1
in two experiments: one ``real'' and one ``simulated.''
In the real experiment (ss-rep-1), $A$ outputs a set $S$ and is given the
$M,\pubaux$ that result from running $\Rep(S)$.
In the simulated experiment (ss-rep-0), $A$ outputs $S$ and is given $M, \pubaux$ output
by a simulator $\calS$ that is only given ${\sf leak}(S)$. \jnote{Note
there is no need for a query oracle here, since the attacker knows the set. Interestingly,
if we provide a query oracle in the real world and an oracle that returns the correct
answer in the simulated world, then this definition seems to imply correctness also.}
For a given function ${\sf leak}$, adversary $A$, and simulator~$\calS$,
we define the advantage measure
$\AdvPrivSS{\Pi,{\sf leak}}{A,\calS} \bydef
\left| \Prob{\ExpPrivSSreal{\Pi}{A}=1} -
\Prob{\ExpPrivSSsim{\Pi,{\sf leak}}{A, \calS}=1} \right|$.
Informally, a scheme $\Pi$ is secure with respect to some leakage ${\sf leak}$
if there is a simulator $\calS$ such that $\AdvPrivSS{\Pi,{\sf leak}}{A,\calS}$ is small
for all~$A$. Intuitively, this means that the public portion
$M, \pubaux$ of the representation of a set~$S$ leaks only ${\sf leak}(S)$.

\begin{figure}[hbtp]
\centering
\fpage{.6}{
\hpagess{.5}{.45}
{
\experimentv{${\bf Exp}^{\rm ss\mbox{-}rep\mbox{-}1}_{\Pi}(A)$}\\
$S \gets A()$ \\
$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
return $A(M,\pubaux)$\\

\medskip
\experimentv{${\bf Exp}^{\rm ss\mbox{-}rep\mbox{-}0}_{\Pi, {\sf leak}}(A, \calS)$}\\
$S \gets A()$\\
$(M,\pubaux) \getsr \calS({\sf info}(S))$ \\
return $A(M,\pubaux)$\\
%\medskip
%\experimentv{$\ExpWPriv{\setprim,\distr{}{}}{A}$}\\
%$S \getsr \distr{}{}$\\
%$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
%$z \getsr A^{\TestOracle}(\pubaux)$\\
%if $z \in S$ then Return 1\\
%Return 0\\
}
%
{}
}
\caption{Experiments for a ``semantic-security'' style privacy
  definition.}
\label{fig:privacy-ss}
\end{figure}

%
%\begin{figure}[htbp]
%\centering
%\hfpages{.495}
%{
%\hpagess{.49}{.49}
%{
%
%\experimentv{$\ExpPrivSSreal{\setprim,\distr{}{}}{A}$}\\
%$S\getsr \distr{}{}$\\
%$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
%$(f,v) \getsr A^{\TestOracle}(M,\pubaux)$\\
%if $f(S)=v$ then\\
%\nudge Return 1\\
%Return 0\\
%
%\medskip
%\experimentv{$\ExpPrivSSsim{\setprim,\distr{}{}}{P}$}\\
%$S\getsr \distr{}{}$\\
%$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
%$(f,v) \getsr P(\pubaux)$\\
%if $f(S)=v$ then\\
%\nudge Return 1\\
%Return 0\\
%}
%%
%{
%\oracle{$\TestOracle(q)$}\\
%$a \gets \Qry(M,\pubaux,\privaux,q)$\\
%Return~$a$\\
%}
%}
%{
%\hpagess{.49}{.49}
%{
%\experimentv{$\ExpPrivSS{\setprim,\distr{}{},P}{A}$}\\
%$S\getsr \distr{}{}$\\
%$b \getsr \bits$\\
%%$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
%$(M_1,\pubaux,\privaux) \getsr \Rep(S)$\\
%$M_0 \getsr P(\pubaux)$\\
%%$(f,v) \getsr \getsr A^{\TestOracle}(M,\pubaux)$\\
%$b' \getsr A^{\TestOracle}(M_b,\pubaux)$\\
%%if $f(S)=v$ then\\
%if $b'=b$ then\\
%\nudge Return 1\\
%Return 0\\
%%
%%\medskip
%%\experimentv{$\ExpPrivSSsim{\setprim,\distr{}{}}{P}$}\\
%%$S\getsr \distr{}{}$\\
%%$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
%%$(f,v) \getsr P(\pubaux)$\\
%%if $f(S)=v$ then\\
%%\nudge Return 1\\
%%Return 0\\
%}
%%
%{
%\oracle{$\TestOracle(q)$}\\
%$a \gets \Qry(M_1,\pubaux,\privaux,q)$\\
%Return~$a$\\
%}
%}
%\caption{Semantic-security style privacy
%  definitions. \textcolor{red}{On the left, need to protect against
%    adversary outputting
%    things like $f(S)=M$.  One idea is to forbid~$f$ such that
%    $f(S)=f(S')$ for all $S,S'$ in the support of $\distr{}{}$.}
%    }
%\label{fig:privacy-ss}
%\end{figure}
